---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# seaice

<!-- badges: start -->
[![R-CMD-check](https://github.com/AustralianAntarcticDivision/seaice/workflows/R-CMD-check/badge.svg)](https://github.com/AustralianAntarcticDivision/seaice/actions)
<!-- badges: end -->

The goal of seaice is to read sea ice concentration data directly from the internet. This package contains functions to 
find the right file URL for a given date, for either northern or southern hemisphere data. 

NOTE: you don't need to use the functions in this package, see the example below which obtains a wrapper function `read_seaice()`. This
package must exist to provide the tools for that to work, but for various reasons I don't want the read function to be in this package.


This package contains a file list (where the file is on the internet, and the date it applies to) for a data file, and will
construct a raster format that can be used to read the data directly. This happens via GDAL, using its virtual raster format "VRT". A temporary file is created to store the information about the file at the URL, and then GDAL does the rest, downloading the file and reading from it into whatever grid/projection we specify. 

With this we can 

- read any date since 1978-10-26 (currently we are a bit limited to the latest but see todo)
- specify a raster grid, of any resolution, any projection, and any extent any where on the planet (the polar regions are really the only relevant ones, but we can use the entire earth or part of it)
- read both hemispheres in one call

## TODO

- currently only NSIDC, add more providers
- keep files up to date, started with 2021-05-23
- be smarter about filling in missing days
- write helpers for commonly used grids (e.g. 0.25 degree longlat, Mercator, or local equal area, etc)
- write helpers for terra, stars, base image(), netcdf output, whatever
- speed it up, currently a bit slow (25s for each file, but subsequent read of each day is fast because the file seems to be 
cached by GDAL or the OS)

These files are binary and read over FTP, so the entire file is downloaded somewhere by GDAL (they are small). Other sources use GeoTIFF or NetCDF and we are exploring including those as well. 



Let us know what you think!  See the [Issues tab](https://github.com/AustralianAntarcticDivision/seaice/issues) to discuss, or post information about any problems you encounter.



## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("AustralianAntarcticDivision/seaice")
```

## Example

This is a basic example which shows you how to read sea ice data for any date since 1978-10-26. The data is projected onto 
a raster grid, with a global default provided. You can use the `xylim` argument to specify a specific raster in any projection. 

Go with the default.

```{r example}
library(seaice)
source(system.file("examples/read_seaice.R", package = "seaice", mustWork = TRUE))
library(raster)  # we could use terra or stars or whatever, see todo 

icecol <- grey.colors(100)
zl <- c(1, 100)
ice0 <- read_seaice("2020-10-15")
plot(ice0, col = icecol, zlim = zl)
```

Usually we will want something more specific, like only the southern hemisphere.  To do that, define the right raster grid
and set the `xylim` argument with it. 

The source data for this function is 25km pixels, in a polar grid. We choose 0.2 in degrees as the resolution (about a fifth of 100km which is about the distance in a degree along a great circle). 

```{r custom}
r <- raster(extent(-180, 180, -72, -55), res = 0.2, crs = "+proj=longlat")
(ice <- read_seaice("2020-10-15", xylim = r))

plot(ice, col = icecol, zlim = zl)
```

Or we can pick our own projection/grid that we want. 

```{r projection}
rg <- raster(extent(-2e6, 2e6, -1e6, 1e6), res = 25000, crs = "+proj=laea +lat_0=-60 +lon_0=180")
pice <- read_seaice("2020-10-15", xylim = rg)
plot(pice, col = icecol, zlim = zl, asp = 1)

## hack a graticule
ll <- rgdal::project(coordinates(rg), projection(rg), inv = TRUE)
ll[ll[,1] < 0, 1] <- ll[ll[,1] < 0, 1] + 360
contour(setValues(rg, ll[,1]), add = TRUE)
contour(setValues(rg, ll[,2]), add = TRUE)

```


To get exactly the grid used by this source data, use a bit of a trick. 

```{r native}
tfile <- nsidc_south_vrt() ## date does not matter
## this is the grid, used by NSIDC, here in the abstract (no data in it)
g <- raster(raster(tfile))

native <- read_seaice("2020-10-15", xylim = g)
plot(native, col = icecol, zlim = zl)
```

This grid is an old one, defined as 'EPSG:3412'. 

```{r hughes}
print(native)
```


# Get involved


Let us know what you think!  See the [Issues tab](https://github.com/AustralianAntarcticDivision/seaice/issues) to discuss, or post information about any problems you encounter.




## Code of Conduct
  
Please note that the seaice project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
